### 2015-05-28 abstract.tex
Bunch of stuff. AAAAAAAAAAAAAAAH
Theory: SSL/TLS.
Comments from Bastien, changed title, added part of the intro.

\thispagestyle{plain}% This way, we can have the page numbering in the foot.
\selectlanguage{english}
% \renewcommand{\abstractname}{\Large{The \lego Motorcyclist robot}}
\addcontentsline{toc}{chapter}{\numberline{}Abstract}
\begin{abstract}
	\begin{verse}
	\small \textit{
	by Quentin Delhaye,  Master in Computer Science and Engineering, Professional Focus, Université Libre de Bruxelles, 2014--2015.}
	\normalsize
	\end{verse}
	\begin{center}
		\textbf{Implementation of high level cryptographic protocol using a SoC platform}
	\end{center}
This is an abstract.
\end{abstract}
\clearpage

\thispagestyle{plain}
\selectlanguage{francais}
% \renewcommand{\abstractname}{\Large{Le robot motard \lego}}
\addcontentsline{toc}{chapter}{\numberline{}Résumé}
\begin{abstract}
	\begin{verse}
	\small \textit{
	par Quentin Delhaye, Master en ingénieur civil en informatique, à finalité spécialisée, Université Libre de Bruxelles, 2014--2015.}
	\normalsize
	\end{verse}
\end{abstract}

>>>>

\thispagestyle{plain}% This way, we can have the page numbering in the foot.
\selectlanguage{english}
\addcontentsline{toc}{chapter}{\numberline{}Abstract}
\begin{abstract}
	\begin{verse}
	\small \textit{
	by Quentin Delhaye,  Master in Computer Science and Engineering, Professional Focus, Université Libre de Bruxelles, 2014--2015.}
	\normalsize
	\end{verse}
	\begin{center}
		\textbf{Implementation of high level cryptographic protocols using a SoC platform}
	\end{center}
This is an abstract. Yep.
\end{abstract}
\clearpage

\thispagestyle{plain}
\selectlanguage{francais}
\addcontentsline{toc}{chapter}{\numberline{}Résumé}
\begin{abstract}
	\begin{verse}
	\small \textit{
	par Quentin Delhaye, Master en ingénieur civil en informatique, à finalité finalité Software and Critical Systems Design, Université Libre de Bruxelles, 2014--2015.}
	\normalsize
	\end{verse}
	\begin{center}
		\textbf{Implementation of high level cryptographic protocols using a SoC platform}
	\end{center}

Abstract, bitch !
\end{abstract}

### 2015-05-28 conclusion.tex
Comments from Bastien, changed title, added part of the intro.

\paragraph{Cryptodev}
If patches adding the GCM support to cryptodev have already been released, those are not compatible with Barco Silex' driver.
Adapting the interface would open the GCM hardware offload to the whole user space applications park.

@@

>>>>

\paragraph{Cryptodev}
If patches adding the GCM support to cryptodev have already been released, those are not compatible with OpenSSL EVP interface.
Adapting the interface would open the GCM hardware offload to the whole user space applications park.

@@

\section{Conclusion}

Security is everywhere.
But security is not cheap, security is power-hungry.

### 2015-05-28 introduction.tex
Some shit.
Comments from Bastien, changed title, added part of the intro.

\chapter{Introduction}
% What is hardware offloading?

@@

Power is critical nowadays, esapcially in datacenters.
Several solutions exists: abandonning general purpose processors altoghether and turn towards ARM cores, or add an FPGA to the processor for example.
Intel proposed the latter solution in 2014~\cite{intel-xeon-fpga} and is on its way to expand that market with the merger with Altera~\cite{intel-altera-merger}.

\section{Challenge}

\section{Network security}

>>>>

\chapter{Introduction}
% What is hardware offloading?
In 1965, Gordon Moore, co-founder of Intel, stated that for the same footprint, the number of transistors in an integrated circuit should double every two years~\cite{moore1965}.
It became the Moore's law, an empirical prediction that drived inovation for fifty years.
Nowadays, it is reaching its limits, as Intel in delaying the manufacturing of its next-generation processors.

The technological limitation is not the only parameter slowing the development of new ICs.
The second law of Moore states that the cost to design, manufacture and test the new architectures doubles every four years.
This economical growth will eventually collide with the Moore's law, stopping its progression.

The domination of multi-purpose processors is reaching its limits, and even the biggest actors on the market are considering alternatives to the Moore's Law.

\noindent A report of the Natural Resources Defense Council~\cite{nrdc2014} forecasts the US datacenter electrity bill to reach \$13.7 billion by 2020.
As CPUs are largely responsible for the power consumption, it makes their design critical.
A solution is to use a hybrid architecture, combining a general-purpose micro-processor with an integrated circuit configurable on-the-fly.
This is a path Intel is exploring with its new Xeon product line, a processor aimed at servers, that combines a Xeon CPU with an FPGA from Altera\cite{xeon2014}.\newline{}

% on its way to expand that market with the merger with Altera~\cite{2015-intel-altera-merger}.

\section{Aim and purpose of this thesis}

% Hardware offloading consists in using a dedicated hardware device to process data instead of the CPU.

In the scope of this work, the testing platform will be the SoCFPGA SoCrates from Altera (see figure~\ref{socrates-photo}.
This SoC combines an ARM dual-core Cortex-A9 with an Altera Cyclone V.
Its ethernet interface will allow us to set the platform in a networked environment and test network security protocols.

This work will focus on the network security and the impact of hardware offloading of cryptographic operations.

@@

Power is critical nowadays, esapcially in datacenters.
Several solutions exists: abandonning general purpose processors altoghether and turn towards ARM cores, or add an FPGA to the processor for example.
Intel proposed the latter solution in 2014~\cite{intel-xeon-fpga} and is on its way to expand that market with the merger with Altera~\cite{2015-intel-altera-merger}.

There a re diffirent ways to add cryptographic capabilities to an application:
\begin{itemize}
	\item Having the cipher built in.
	\item Using a cryptographic library, such as OpenSSL.
	\item Using the Linux Crypto API.
\end{itemize}

Present the two IP cores that will be used in this work: BA411E and BA414E.

\begin{figure}[ht]
\includegraphics[width=\linewidth]{socrates-photo}
\caption{Altera SoCrates platform}{Combines an ARM Cortex-A9 with an Altera Cyclone V.}
\label{socrates-photo}
\end{figure}

\section{Context}
The work presented in this thesis has been conducted under the supervision of Sébastien Rabou, in Barco Silex' office in Louvain-la-Neuve.

\section{Structure}
The present document will begin with the technical ground necessary to this work.
After a presentation on how to drive a hardware device from the operating system and details on device driver design, we will adress the cryptography.
Only most relevant part of cryptography will be detailed: message digests, symmetric cryptography and more specifically the CBC and GCM modes of operation, and asymmetric cryptography with an accent on RSA and Diffie-Hellman.

The next chapter will present the implementation methods for the protocols theoretically introduced.
Four applications are considered: OpenVPN, OpenSSH, OpenSSL and Strongswan.

Before presenting the results of those implementations, the tests protocols will be discussed and the platforms introduced.
Three tests will be undertaken: TLS connections, latency and file transfer.

The results will address those three tests for different implementations, and the impact of hardware offloading will be analyzed and compared.

Finally, some future works and directions that the project could take are gathered in the last chapter, followed by a concluding word.

### 2015-05-28 results.tex
Comments from Bastien, changed title, added part of the intro.

As for the combination of encryption and authentication, IPsec is between 13\% and 32\% faster than OpenVPN.
IPsec loses its advantage with the increase of the payload size, the time lost when moving around the data between the user and the kernel space being less important compared to the securoty operations.

>>>>

As for the combination of encryption and authentication, IPsec is between 13\% and 32\% faster than OpenVPN.
IPsec loses its advantage with the increase of the payload size, the time lost when moving around the data between the user and the kernel space being less important compared to the security operations.

### 2015-05-28 technical-background.tex
Comments from Bastien, changed title, added part of the intro.

In order for the operating system to comunicate with the device, they will need to share some memory.
That memory can be directly mapped and accessed from the user-space using \texttt{/dev/mem}, or can use a direct memory access module (DMA).
From the operating system, we build a bunch of scatterlist in the kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
They can be mapped three different ways: \texttt{DMA\_BIDIRECTIONAL}, \texttt{DMA\_TO\_DEVICE} or \texttt{DMA\_FROM\_DEVICE}.
When the CPU write something in those descriptors and synchronize them with the DMA, it does not have to care about them anymore, the DMA in now in charge to send them to the device where registers are ready to read the incomming data.
The same goes from the device to the CPU: when the device wants to communicate data to the OS, it writes it on the DMA that will transfer them to the CPU, triggering a flag on the way to notify it.

@@

There are two types of interface with such a device: master or slave.
The figure~\ref{fig:barco-ba411e-ipcore}  represent one of Barco Silex' crypto IP cores that proposes two interfaces: one of each type.
\begin{itemize}
	\item The data path is a master. That is the operating system will not have a direct access to the memory registers of the hardware devide. Instead, they will communicate via a DMA bus: one peer places data on the DMA and notifies the other that he can fetch it.
	\item The control (ctrl) path is a slave. The operatin system can directly read and write on this part of the device memory.
\end{itemize}

A master interface has the advantage to offload a part of the memory management to the device, but there is a space overhead to connect the DMA as well as an extra delay to process its data.
A slave interface is usually used for sporadic data, avoiding the overhead of a DMA.
As an example, the data path of Barco Silex' symmetric cryptography IP core is a master, as there is a lot of data to process, whilst the same data path of their asymmetric cryptography IP core is a slave, for the operands are larger and the computation takes more time.

@@

There are two ways to get the status of a hardware device: either by using interruption, or by actively polling the device, relentlessly asking if it finished its operations.
The first case is the cleanest and the most common: when the device has something to send to the driver, or if anything unexpected happened, it sends an interruption request (IRQ) to the processor, which will in turn execute the interruption service routine (ISR) registered by the driver~\citep[chap. 10]{Corbet:2005:LDD:1209083}.
The second case is the easiest and is always guaranteed to work, but won't let go of the processor willingly, loading it at 100\%, and avoiding any other task to be executed.
Fortunately, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, that is the scheduler interrupts the running task and assigns the processor ressources it used to an other one.
Hence, systems with a lot of processes in need for CPU ressources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.\newline{}

The vast majority of drivers are meant to be run in kernel space, but some use cases may require the user space to directly access the hardware.
In those cases, one can use the Userspace I/O framework.

@@

There exists numerous cipher types, and one of the most widespread is the block cipher.
A block cipher splits the message into blocks of fixed size and work on them separately.
An example of such a cipher is illustrated on the figure~\ref{fig:cbc-encrypt} with the Cipher Block Chaining (CBC) mode.
This particular mode is easy to implement and has been robust for a long time, before the rise of the padding oracle attack~\cite{vaudenay2002}.

\indent An other interresting block cipher mode is the Counter (CTR).
Each block of data is treated independently from the others (see figure~\ref{fig:ctr-encrypt}, making the scheme parallelizable and thus more interresting on hardware.

@@

\noindent New block cipher have emerged to answer the need for such combinations and formed the Authenticated Encryption with Associated Data (AEAD) mode of operation.
The most widespread mode in network applications is the Galois Counter Mode (GCM, see figure~\ref{fig:gcm-encrypt}), combining the CTR block cipher mode and a GHASH function using multiplications in the Galois field.
The advantage of this mode is that the authentication can be computed in parallel with the encryption, itself parallelizable.
This mode can thus be highly optimized in hardware, as proved Barco Silex with their BA415 IP core reaching up to 100Gbps~\cite{barco-ba415}.

>>>>

The operating system and the device need to share some memory to communicate.
One possibility is to directly map and access from the user-space using \texttt{/dev/mem}, and an other is to use a direct memory access module (DMA).

@@

There are two types of interface with such a device: master or slave.
The figure~\ref{fig:barco-ba411e-ipcore}  represent one of Barco Silex' crypto IP cores that proposes two interfaces: one of each type.
\begin{itemize}
	\item The data path is a master. That is the operating system will not have a direct access to the data registers of the hardware device. Instead, they will communicate via a DMA: one peer places data on the DDR memory and notifies the other that he can fetch it.
	\item The control (ctrl) path is a slave. The operating system can directly read and write on this part of the device memory.
\end{itemize}

A master interface has the advantage to offload data access operations to the device, but there is a ressource and area overhead to connect the DMA as well as an extra complexity to process its data.
A slave interface is usually used for sporadic data (\textit{e.g.} status, control flags), avoiding the overhead of a DMA.
As an example, the data path of Barco Silex' symmetric cryptography IP core is a master, as there is a lot of data to process, whilst the same data path of their asymmetric cryptography IP core is a slave, for the operands are larger and the computation takes more time.\newline{}

If a master interface with a DMA is used, the operating system buils scatterlists in the kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
They can be mapped three different ways: \texttt{DMA\_BIDIRECTIONAL}, \texttt{DMA\_TO\_DEVICE} or \texttt{DMA\_FROM\_DEVICE}.
When the CPU write something in those descriptors and synchronize them with the DMA, it does not have to care about them anymore, the DMA in now in charge to send them to the device where registers are ready to read the incomming data.
The same goes from the device to the CPU: when the device wants to communicate data to the OS, it writes it on the DMA that will transfer them to the CPU, triggering a flag on the way to notify it.

@@

There are two ways to get the status of a hardware device: either by using interruption, or by actively polling the status register fo the device, relentlessly asking if it finished its operations or tasks.
The first case is the cleanest and the most common: when the hardware device has something to send to the driver, or if anything unexpected happened, it sends an interruption request (IRQ) to the processor, which will execute the interruption service routine (ISR) registered by the driver~\citep[chap. 10]{Corbet:2005:LDD:1209083}.
The second case is the easiest and is always guaranteed to work, but will monopolize the processir, loading it at 100\%, and avoiding any other task to be executed.
Fortunately, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, \textit{i.e.} the scheduler interrupts the running task and assigns the processor ressources it used to an other one.
Hence, systems with a lot of processes in need for CPU ressources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.\newline{}

The vast majority of drivers are meant to be run in kernel space, but some use cases may require the user space to directly access the hardware.
In those cases, one can use the Userspace I/O framework on a slave interface.

@@

There exists numerous cipher types, and one of the most widespread is the block cipher.
A block cipher splits the message into blocks of fixed size and work on them separately.
An example of such a cipher is illustrated on the figure~\ref{fig:cbc-encrypt} with the Cipher Block Chaining (CBC) mode.
This particular mode is easy to implement and has been considered robust for a long time, before the rise of the padding oracle attack~\cite{vaudenay2002}.

\indent An other interresting block cipher mode is the Counter (CTR).
Each block of data is treated independently from the others (see figure~\ref{fig:ctr-encrypt}), making the scheme parallelizable and thus more interresting on hardware.

@@

\noindent New block cipher have emerged to answer the need for such combinations and formed the Authenticated Encryption with Associated Data (AEAD) mode of operation.
The most widespread mode in network applications is the Galois Counter Mode (GCM, see figure~\ref{fig:gcm-encrypt}), combining the CTR block cipher mode and a GHASH function using multiplications in the Galois field.
The advantage of this mode is that the authentication can be computed in parallel with the encryption, itself parallelizable.
This mode can thus be highly optimized in hardware, as proved by Barco Silex with their BA415 IP core reaching up to 100Gbps~\cite{barco-ba415}.
